---
title: "The powdR package"
output: 
  html_document:
    includes:
      after_body: footer.html
bibliography: bibliography.bib
---

  <style type="text/css">
  .title {
    display: none;
  }
  
  body{ /* Normal  */
      font-size: 16px;
  }
  code.r{ /* Code block */
    font-size: 16px;
  }
  p {line-height: 1.5em;}
  
  }
  </style>

<div class="row" style="padding-top: 30px;">
<div class="col-sm-9">


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# **The `powdR` package**

[![Build Status](https://travis-ci.org/benmbutler/powdR.svg?branch=master)](https://travis-ci.org/benmbutler/powdR)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/powdR)](https://CRAN.R-project.org/package=powdR)
[![Total_Downloads](https://cranlogs.r-pkg.org/badges/grand-total/powdR)](https://cran.r-project.org/package=powdR)

---

## Overview

`powdR` is an open source implementation of the full pattern summation approach to quantitative mineralogy from X-ray powder diffraction data [@Chipera2002;@Chipera2013;@Eberl2003], written in the `R` programming language. It seeks to make quantitative phase analysis from XRPD data a faster and more user-friendly procedure.

This short tutorial explains the installation and basic use of `powdR`, which requires a prior installation of R. A more detailed tutorial of the package is provided in the pacakge [Vignette](https://cran.r-project.org/web/packages/powdR/vignettes/powdR_vignette.html), and video tutorials of the `powdR` web application can be found on the powdR [YouTube channel](https://www.youtube.com/channel/UC4na-XpeEoTU7rB5aMpTQHw).

## Installation

The stable version of `powdR` is on CRAN:
```{r, eval = FALSE}
install.packages("powdR")
```

Alternatively, the development version can be downloaded from GitHub
```{r, eval = FALSE}
#Install devtools if you don't already have it on your machine
install.packages("devtools")

devtools::install_github('benmbutler/powdR')
```

## Basic usage

```{r example}
library(powdR)

#Load some soils to quantify
data(soils)

#Load a powdRlib reference library of pure patterns
data(minerals)

#Quantify a sample
q <-  fps(lib = minerals,
          smpl = soils$sandstone,
          refs = minerals$phases$phase_id,
          std = "QUA.1")

#Inspect the phase concentrations (summarised by name)
q$phases_grouped

#Inspect the quantification
plot(q, wavelength = "Cu")

```

Alternatively, `plot(q, wavelength = "Cu", interactive = TRUE)` provides an interactive plot for better inspection of the fit.

## The `powdR` Shiny app

To run `powdR` via the Shiny web applicationp, use `run_powdR()`. This loads the application in your default web browser. The application has eight tabs:

1. **Reference Library Builder:** Allows you to create and export a `powdRlib` reference library from two .csv files: one for the XRPD measurements, and the other for the ID, name and reference intensity ratio of each pattern.
2. **Reference Library Viewer:** Facilitates quick inspection of the phases within a `powdRlib` reference library.
3. **Reference Library Editor:** Allows the user to easily subset a `powdRlib` reference library .
4. **Full Pattern Summation:** A user friendly interface for iterative full pattern summation of single samples.
5. **Automated Full Pattern Summation:** A user friendly interface for automated full pattern summation of single samples.
6. **Results viewer:** Allows for quick inspection of results derived from full pattern summation.
7. **Results editor:** Allows for results from previously saved `powdRfps` and `powdRafps` objects to be edited via addition or removal of reference patterns to the fitting process.
8. **Help** Provides a series of video tutorials (via YouTube) detailing the use of the `powdR` Shiny application.

## References



